Windowed least-squares model reduction for dynamical systems

### 2019-10-18 section3/space_time.tex

%Inserting the approximation~\eqref{eq:stapprox} into the optimization problem~\eqref{eq:tclsrm}, the
%constrained minimization problem over the $n$th time slab can be re-written as,
%\begin{equation}\label{eq:obj_gen_slab_spacetime} \stgenstate^n =
%\underset{\stgenstatey}{\text{argmin }} \mathcal{J}^n \bigg( \stbasisArgnt{n} \stgenstatey + \stateInterceptArg{n}  \bigg) , \end{equation}
%%\begin{equation}\label{eq:obj_gen_slab} \genstate^n(t) =
%%\underset{\genstatey}{\text{argmin }} \mathcal{J}^n
%%\bigg(\decoder\big(\genstatey(\tau)\big) \bigg) , \end{equation}
%subject to the boundary conditions 
%\begin{equation}\label{eq:bcs_spacetime}
%\stbasisArg{n}{\timeStartArg{n}}\stgenstateArg{n} = \begin{cases} 
%\mathbb{P}^n\stbasisArg{n-1}{\timeEndArg{n-1}}\stgenstateArg{n-1}  & n = 2,\ldots,\nslabs, \\ 
%\mathbb{P}^n \stateFOMIC
% & n=1. \end{cases} \end{equation}
\subsubsection{Stationary conditions} 
The key difference between \spaceTimeAcronym\ and \spatialAcronym\ trial
	subspaces is as follows: generalized coordinates for \spaceTimeAcronym\ trial subspaces 
	comprise a vector in $\RR{\stdimArg{n}}$, while generalized coordinates for
	the \spatialAcronym\ trial subspaces comprise a
	time-dependent vector in
	$\RR{\romdimArg{n}} \otimes \timeSpaceArg{n}$. 
Thus, in the \spaceTimeAcronym\ case, the optimization problem is no longer
	minimizing a functional with respect to a (time-dependent) function, but now is minimizing a function with 
respect to a vector.  As such, the first-order optimality conditions can be
	derived using standard calculus. Differentiating the objective with respect
	to the generalized coordinates and setting the result equal to zero yields 
\begin{equation}\label{eq:st_stationary}
 \intSlabArg{n} \bigg[ \stbasisDotArg{n}{t}^T  - \stbasisArg{n}{t}^T \bigg[\frac{\partial
\velocity}{\partial \stateyDiscrete} (\stbasisDotArg{n}{t} \stgenstateArg{n} +                    
\stateInterceptSTArg{n},t)\bigg]^T  \bigg] \stweightingMatArg{n} \bigg( \stbasisDotArg{n}{t} \stgenstateArg{n}  - \velocity (\stbasisArg{n}{t} \stgenstateArg{n} + \stateInterceptSTArg{n},t) \bigg) dt = \bz.\end{equation}
%subject to the boundary conditions 
%\begin{equation}\label{eq:bcs_spacetime}
%\stbasisArg{n}{\timeStartArg{n}}\stgenstateArg{n} = \begin{cases} 
%\mathbb{P}^n\stbasisArg{n-1}{\timeEndArg{n-1}}\stgenstateArg{n-1}  & n = 2,\ldots,\nslabs, \\ 
%\mathbb{P}^n \stateFOMIC &
%n=1. \end{cases} \end{equation}
Thus, for \spaceTimeAcronym\ trial subspaces, the stationary conditions
	comprise a system of algebraic equations, as opposed to a system of
	differential equations as with \spatialAcronym\ trial subspaces.

>>>>

%Inserting the approximation~\eqref{eq:stapprox} into the optimization problem~\eqref{eq:tclsrm}, the
%constrained minimization problem over the $n$th time slab can be re-written as,
%\begin{equation}\label{eq:obj_gen_slab_spacetime} \stgenstate^n =
%\underset{\stgenstatey}{\text{argmin }} \mathcal{J}^n \bigg( \stbasisArgnt{n} \stgenstatey + \stateInterceptArg{n}  \bigg) , \end{equation}
%%\begin{equation}\label{eq:obj_gen_slab} \genstate^n(t) =
%%\underset{\genstatey}{\text{argmin }} \mathcal{J}^n
%%\bigg(\decoder\big(\genstatey(\tau)\big) \bigg) , \end{equation}
%subject to the boundary conditions 
%\begin{equation}\label{eq:bcs_spacetime}
%\stbasisArg{n}{\timeStartArg{n}}\stgenstateArg{n} = \begin{cases} 
%\mathbb{P}^n\stbasisArg{n-1}{\timeEndArg{n-1}}\stgenstateArg{n-1}  & n = 2,\ldots,\nslabs, \\ 
%\mathbb{P}^n \stateFOMIC
% & n=1. \end{cases} \end{equation}
\subsubsection{Stationary conditions} 
The key difference between \spaceTimeAcronym\ and \spatialAcronym\ trial
	subspaces is as follows: generalized coordinates for \spaceTimeAcronym\ trial subspaces 
	comprise a vector in $\RR{\stdimArg{n}}$, while generalized coordinates for
	the \spatialAcronym\ trial subspaces comprise a
	time-dependent vector in
	$\RR{\romdimArg{n}} \otimes \timeSpaceArg{n}$. 
Thus, in the \spaceTimeAcronym\ case, the optimization problem is no longer
	minimizing a functional with respect to a (time-dependent) function, but is minimizing a function with 
respect to a vector.  As such, the first-order optimality conditions can be
	derived using standard calculus. Differentiating the objective function with respect
	to the generalized coordinates and setting the result equal to zero yields 
\begin{equation}\label{eq:st_stationary}
 \intSlabArg{n} \bigg[ \stbasisDotArg{n}{t}^T  - \stbasisArg{n}{t}^T \bigg[\frac{\partial
\velocity}{\partial \stateyDiscrete} (\stbasisDotArg{n}{t} \stgenstateArg{n} +                    
\stateInterceptSTArg{n},t)\bigg]^T  \bigg] \stweightingMatArg{n} \bigg( \stbasisDotArg{n}{t} \stgenstateArg{n}  - \velocity (\stbasisArg{n}{t} \stgenstateArg{n} + \stateInterceptSTArg{n},t) \bigg) dt = \bz.\end{equation}
%subject to the boundary conditions 
%\begin{equation}\label{eq:bcs_spacetime}
%\stbasisArg{n}{\timeStartArg{n}}\stgenstateArg{n} = \begin{cases} 
%\mathbb{P}^n\stbasisArg{n-1}{\timeEndArg{n-1}}\stgenstateArg{n-1}  & n = 2,\ldots,\nslabs, \\ 
%\mathbb{P}^n \stateFOMIC &
%n=1. \end{cases} \end{equation}
Thus, for \spaceTimeAcronym\ trial subspaces, the stationary conditions
	comprise a system of algebraic equations, as opposed to a system of
	differential equations as with \spatialAcronym\ trial subspaces.

### 2019-10-18 section4/numerical_methods.tex

\section{Numerical solution techniques}\label{sec:numerical_techniques}
We now consider numerical-solution techniques for the \methodAcronym\ \approachKwd.
We focus primarily on 
direct (i.e., discretize-then-optimize) and indirect (i.e., optimize-then-discretize) methods for \spatialAcronym\ trial subspaces, 
and then briefly outline direct and indirect methods for \spaceTimeAcronym\ trial subspaces. 
%It is noted that, as \methodAcronym\ with \spatialAcronym\ trial subspaces 
%Given the formulation of the \methodAcronym\ method presented in Section
%\ref{sec:tclspg}, we now consider numerical-solution techniques for the method. 
%Due to the commonalities that the \methodAcronym\ method exhibits with
%optimal-control problems, \KTC{is this only for S-reduction subspaces? If so,
%state this.} the derivation of these techniques can leverage a large existing
%body of work. In this section, we outline a variety of such techniques. 
\subsection{\spatialAcronym: direct and indirect methods}
\methodAcronym\ with \spatialAcronym\ trial subspaces can be viewed as an optimal-control problem.
Numerical techniques to
solve optimal-control problems can be classified as either
\textit{direct} or \textit{indirect}
methods~\cite{conway_optimalcontrolreview}. Rather than working with the first order optimality conditions~\eqref{eq:lspg_continuous}--\eqref{eq:lspg_bcs}, direct methods ``directly" solve the optimization problem~\eqref{eq:tclsrm} by first
numerically discretizing the system and objective functional and ``transcribing"
the infinite dimensional problem into a finite-dimensional nonlinear
programming (optimization) problem. Direct approaches thus
``discretize-then-optimize''.
In contrast, indirect methods compute solutions to the Euler--Lagrange equations~\eqref{eq:lspg_continuous}--\eqref{eq:lspg_bcs} (which comprise the first-order optimality conditions of the objective functional). Thus, indirect methods ``optimize-then-discretize," and solve the optimization problem
``indirectly". A variety of both discretization and solution techniques are possible
for both direct and indirect methods. Collocation methods,
finite-element methods, spectral methods, and shooting methods are all examples of
plausible solution techniques.

>>>>

\section{Numerical solution techniques}\label{sec:numerical_techniques}
We now consider numerical-solution techniques for the \methodAcronym\ \approachKwd.
We focus primarily on 
direct (i.e., discretize-then-optimize) and indirect (i.e., optimize-then-discretize) methods for \spatialAcronym\ trial subspaces, 
and then briefly outline direct and indirect methods for \spaceTimeAcronym\ trial subspaces. 
%It is noted that, as \methodAcronym\ with \spatialAcronym\ trial subspaces 
%Given the formulation of the \methodAcronym\ method presented in Section
%\ref{sec:tclspg}, we now consider numerical-solution techniques for the method. 
%Due to the commonalities that the \methodAcronym\ method exhibits with
%optimal-control problems, \KTC{is this only for S-reduction subspaces? If so,
%state this.} the derivation of these techniques can leverage a large existing
%body of work. In this section, we outline a variety of such techniques. 
\subsection{\spatialAcronym: direct and indirect methods}
\methodAcronym\ with \spatialAcronym\ trial subspaces can be viewed as an optimal-control problem.
Numerical solution techniques for
optimal-control problems can be classified as either
\textit{direct} or \textit{indirect}
methods~\cite{conway_optimalcontrolreview}. Rather than working with the first-order optimality conditions~\eqref{eq:lspg_continuous}--\eqref{eq:lspg_bcs}, direct methods ``directly" solve the optimization problem~\eqref{eq:tclsrm} by first
numerically discretizing the system \KTC{What does it mean to discretize the
system?} and objective functional and ``transcribing"
the infinite-dimensional problem into a finite-dimensional 
problem. Direct approaches thus
``discretize then optimize''. \KTC{Note: only use hyphens when
discretize-then-optimize is used as an adjective.}
In contrast, indirect methods compute solutions to the Euler--Lagrange
equations~\eqref{eq:lspg_continuous}--\eqref{eq:lspg_bcs} (which comprise the
first-order optimality conditions). \KTC{Note: they are not first-order
optimality conditions of the objective functional only; they are first-order
optimality conditions of the optimization problem} Thus, indirect methods
``optimize then discretize," and solve the optimization problem
``indirectly.'' A variety of both discretization and solution techniques are possible
for both direct and indirect methods. Collocation methods,
finite-element methods, spectral methods, and shooting methods are all examples of
possible solution techniques.

### 2019-10-18 section4/spatial_numerical.tex

\methodAcronym\ with the direct approach and a linear multistep method sequentially computes the solutions
$\approxstateDiscreteArg{n,1},\ldots,\approxstateDiscreteArg{n,\nstepsArg{n}}$,
$n=1,\ldots,\nslabs$, that satisfy
%	\KTC{What is $\genstate^{n-1}(\timeEndArg{n-1} ) $ doing here? Two problems
%	(1) there is no notion of a generalized coordinate $\genstate$ yet, and (2)
%	this should be the discrete solution, not the exact solution evaluated at
%	the final time.} \EP{You are correct, I think there is an error now in the minimize term though. It was $\in \trialspace\otimes \RR{\lmsWidthArg{n}{i}+1}$, it should be $\trialspace\otimes \RR{\nstepsArg{n}+1}$}
\begin{align*}
	&\underset{(\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}) \in ( \trialspace + \stateInterceptArg{n}) \otimes \RR{\nstepsArg{n}}}{\text{minimize } }
\objectiveArgLMS{n}_{\text{D-S}} (\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}), 
%%& \text{subject to} \hspace{0.2 in}  \stateyDiscreteArg{0} =
%%\begin{cases} \basisspaceArg{n} \spatialICDiscrete + \stateInterceptArg{n}& n = 2,\ldots,\nslabs,\\
%\basisspaceArg{1} \genstateICOne  + \stateInterceptArg{1}& n=1. \end{cases} 
\end{align*}
where
\KTC{This next equation: I don't think you should project the previous
solutions.}
\begin{equation*}
\begin{split}
\objectiveArgLMS{n}_{\text{D-S}} &\vcentcolon  (\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}) \mapsto \objectiveArgLMS{n}(\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}; \projectorSpatialArg{n}(\approxstateDiscreteArg{\indexMapper(n,0)}),\ldots,\projectorSpatialArg{n}(\approxstateDiscreteArg{\indexMapper(n,-\lmsWidthArg{n}{1}+1)})), \\
&\vcentcolon \RR{\fomdim} \otimes \RR{\nstepsArg{n}} \rightarrow \RR{}_+,
\end{split}
\end{equation*}
and $\projectorSpatialArg{n} : \RR{\fomdim} \rightarrow \trialspaceArg{n}$ is the $\elltwo$-orthogonal projector 
$$\projectorSpatialArg{n} : \stateyDiscreteArgnt{} \mapsto  \basisspaceArg{n}\basisspaceTArg{n} (\stateyDiscreteArgnt{} - \stateInterceptArg{n}) + \stateInterceptArg{n} .$$
We additionally set $\approxstateDiscreteArg{1,0} = \projectorSpatialArg{n}(\stateFOMIC).
$
%where 
%\begin{align*}
%\residArg{n,i} &: (\stateyArgnt{n,i},\ldots,\stateyArgnt{n,i-k(i)}) \mapsto  \sum_{j=0}^k \alpha_j \stateyArgnt{n,i-j} = \Delta t^{n,i} \sum_{j=0}^k \beta_j \velocity(\stateyArgnt{n,i-j}),
%\\
%               &: \RR{\fomdim} \otimes \RR{k(i)} \rightarrow \RR{\fomdim}, 
%\end{align*} 
%is the discrete residual resulting from the linear multistep scheme and 
%\methodAcronym\ solves the minimization problem over each time window, 
%\begin{align*}
%&\approxstateDiscreteArg{n,0},\ldots,\approxstateDiscreteArg{n,\nstepsArg{n}} = \underset{\stateyDiscreteArgnt{n,0},\ldots,\stateyDiscreteArgnt{n,\nstepsArg{n}} \in \trialspace}{\text{arg\,min } }
%\objectiveArgLMS{n} (\stateyDiscreteArgnt{n,0},\ldots,\stateyDiscreteArgnt{n,\nstepsArg{n}}) \\
%& \text{subject to} \hspace{0.2 in}  \approxstateDiscreteArg{n,0} =
%\begin{cases} \approxstateDiscreteArg{n-1,\nstepsArg{n-1}} & n = 2,\ldots,\nslabs,\\
%\approxstateIC & n=1. \end{cases} \end{align*}

>>>>

\methodAcronym\ with the direct approach and a linear multistep method sequentially computes the solutions
$\approxstateDiscreteArg{n,1},\ldots,\approxstateDiscreteArg{n,\nstepsArg{n}}$,
$n=1,\ldots,\nslabs$, that satisfy
%	\KTC{What is $\genstate^{n-1}(\timeEndArg{n-1} ) $ doing here? Two problems
%	(1) there is no notion of a generalized coordinate $\genstate$ yet, and (2)
%	this should be the discrete solution, not the exact solution evaluated at
%	the final time.} \EP{You are correct, I think there is an error now in the minimize term though. It was $\in \trialspace\otimes \RR{\lmsWidthArg{n}{i}+1}$, it should be $\trialspace\otimes \RR{\nstepsArg{n}+1}$}
\begin{align*}
	&\underset{(\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}) \in ( \trialspace + \stateInterceptArg{n}) \otimes \RR{\nstepsArg{n}}}{\text{minimize } }
\objectiveArgLMS{n}_{\text{D-S}} (\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}), 
%%& \text{subject to} \hspace{0.2 in}  \stateyDiscreteArg{0} =
%%\begin{cases} \basisspaceArg{n} \spatialICDiscrete + \stateInterceptArg{n}& n = 2,\ldots,\nslabs,\\
%\basisspaceArg{1} \genstateICOne  + \stateInterceptArg{1}& n=1. \end{cases} 
\end{align*}
where
\KTC{This next equation: I don't think you should project the previous
solutions.}
\begin{equation*}
\begin{split}
\objectiveArgLMS{n}_{\text{D-S}} &\vcentcolon  (\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}) \mapsto \objectiveArgLMS{n}(\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}; \approxstateDiscreteArg{\indexMapper(n,0)},\ldots,\approxstateDiscreteArg{\indexMapper(n,-\lmsWidthArg{n}{1}+1)}), \\
&\vcentcolon \RR{\fomdim} \otimes \RR{\nstepsArg{n}} \rightarrow \RR{}_+,
\end{split}
\end{equation*}
with the initial condition $\approxstateDiscreteArg{1,0} = \basisspaceArg{n}\basisspaceTArg{n} (\stateFOMIC - \stateInterceptArg{n}) + \stateInterceptArg{n}$.
%We additionally set $\approxstateDiscreteArg{1,0} = \projectorSpatialArg{n}(\stateFOMIC)$, where
%$\projectorSpatialArg{n} : \RR{\fomdim} \rightarrow \trialspaceArg{n}$ is the $\elltwo$-orthogonal projector 
%$$\projectorSpatialArg{n} : \stateyDiscreteArgnt{} \mapsto  \basisspaceArg{n}\basisspaceTArg{n} (\stateyDiscreteArgnt{} - \stateInterceptArg{n}) + \stateInterceptArg{n} .$$

%where 
%\begin{align*}
%\residArg{n,i} &: (\stateyArgnt{n,i},\ldots,\stateyArgnt{n,i-k(i)}) \mapsto  \sum_{j=0}^k \alpha_j \stateyArgnt{n,i-j} = \Delta t^{n,i} \sum_{j=0}^k \beta_j \velocity(\stateyArgnt{n,i-j}),
%\\
%               &: \RR{\fomdim} \otimes \RR{k(i)} \rightarrow \RR{\fomdim}, 
%\end{align*} 
%is the discrete residual resulting from the linear multistep scheme and 
%\methodAcronym\ solves the minimization problem over each time window, 
%\begin{align*}
%&\approxstateDiscreteArg{n,0},\ldots,\approxstateDiscreteArg{n,\nstepsArg{n}} = \underset{\stateyDiscreteArgnt{n,0},\ldots,\stateyDiscreteArgnt{n,\nstepsArg{n}} \in \trialspace}{\text{arg\,min } }
%\objectiveArgLMS{n} (\stateyDiscreteArgnt{n,0},\ldots,\stateyDiscreteArgnt{n,\nstepsArg{n}}) \\
%& \text{subject to} \hspace{0.2 in}  \approxstateDiscreteArg{n,0} =
%\begin{cases} \approxstateDiscreteArg{n-1,\nstepsArg{n-1}} & n = 2,\ldots,\nslabs,\\
%\approxstateIC & n=1. \end{cases} \end{align*}
